---
title: "Problem Set 2"
author: "Sravan Ramaswamy"
date: "2/7/2021"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=TRUE}
#Clear Environment 
rm(list = ls())

#Load Libraries 
library(MASS)
library(ggplot2)
library(evd)
library(dplyr)
library(patchwork)
library(statar)
library(stats)
library(splines)
library(caret)
library(binsreg)
```

```{r setupcont, include=TRUE}
#Load Bid Data
DataDir = paste(getwd(), "/nodelmp.csv",sep = "")
Data = read.csv(DataDir)

```


## Question 1

```{r Q1, include=TRUE}
#Obtain the Maximum Price each day and the corresponding temperature
lmp_Maximum_df <- subset(Data, hrank == 1)

#Make a binned scatterplot of the relationship
lmp_max <- lmp_Maximum_df$lmp
temperature <- lmp_Maximum_df$temp

ggplot(lmp_Maximum_df, aes(x=temp,y=lmp)) +
  stat_smooth(method = "lm", formula = y~x, size = 1, se = FALSE) +
   stat_binmean(n = 10)


```

Based on the binned scatterplot, the relationship between the maximum price per day and the temperature does not appear to be linear.


## Question 2: 


```{r}


rmse <- function(sm) 
    sqrt(mean(sm$residuals^2))

fit <- lm( lmp_max~temperature)
deg = 0

for(i in 1:10){
  fit_curr <- lm(lmp_max~poly(temperature,i))
  if(rmse(fit_curr) < rmse(fit)){
    fit <- fit_curr
    deg = i
  }
}

deg
rmse(fit)

```

According to the mean squared error, the polynomial with a degree of 10 features the best fit in-sample.

## Question 3


```{r}

kmse <- function(x){
  # Define training control
  set.seed(123) 
  train.control <- trainControl(method = "cv", number = 10)
# Train the model
  if(x == 1){
  model <- train(lmp_max ~ temperature, data = data.frame(lmp_max, temperature), method = "lm",
                trControl = train.control)
  }
  else{
    f <- bquote(lmp_max ~ poly(temperature, .(x)))
  model <- train(as.formula(f), data = data.frame(lmp_max, temperature), method = "lm",
                trControl = train.control)
  }
    
  # Summarize the results
  return(model)
}


model <- kmse(1)
deg = 1
for(i in 2:10){
  model_curr <- kmse(i)
  if(model_curr$results$RMSE < model$results$RMSE){
    model <- model_curr
    deg = i
  }
}
deg
model


```

According to the 10-fold cross-validation, the polynomial with the best fit is of degree 7.

## Question 4

```{r}

ggplot(lmp_Maximum_df, aes(x=temp,y=lmp)) +
   stat_smooth(method = "lm", formula = y~x, size = 1, se = FALSE) +
  stat_smooth(method = "lm", formula = y~poly(x,2), size = 1, se = FALSE, color = "red") +
  stat_smooth(method = "lm", formula = y~poly(x,7), size = 1, se = FALSE, color = "black") +
 stat_smooth(method = "lm", formula = y~poly(x,10), size = 1, se = FALSE, color = "orange") +
  stat_binmean(n = 10)

```

## Question 5

```{r}
lmp_Maximum_df <- lmp_Maximum_df[order(lmp_Maximum_df$temp),]

ncsmse <- function(x){
  # Define training control
  set.seed(123) 
  train.control <- trainControl(method = "cv", number = 10)
# Train the model
  f <- bquote(lmp_max ~ ns(temperature, .(df = x + 1)))
  model <- train(as.formula(f), data = data.frame(lmp_max, temperature), method = "lm",
                trControl = train.control)
    
  # Return the results
  return(model)
}

modelspline <- ncsmse(1)
knots = 1
for(i in 2:10){
  model_curr <- ncsmse(i)
  if(model_curr$results$RMSE < model$results$RMSE){
    modelspline <- model_curr
    knots = i
  }
}
knots
modelspline

```
The Optimal number of knots here is 4 knots.

## Question 6


```{r}

fm1 <- loess(lmp_max ~ temperature)
fm1

ggplot(lmp_Maximum_df, aes(x=temp,y=lmp)) +
  stat_smooth(method = "lm", formula = y~poly(x,7), size = 1, se = FALSE, color = "black") +
  stat_smooth(method = "lm", formula = y~ns(x, df = 4+1), size = 1, se = FALSE, color = "blue") +
 stat_smooth(method = "loess", formula = y~x, size = 1, se = FALSE, color = "orange") +
  stat_binmean(n = 10)

```

## Question 7

```{r}


loessresiduals <- fm1$residuals

lmp_Maximum_df$polypredicted <- predict(model$finalModel)
lmp_Maximum_df$polyresiduals <- residuals(model$finalModel)

lmp_Maximum_df$splinepredicted <- predict(modelspline$finalModel)
lmp_Maximum_df$splineresiduals <- residuals( modelspline$finalModel)

lmp_Maximum_df$loesspredicted <- fm1$fitted
lmp_Maximum_df$loessresiduals <- fm1$residuals

ggplot(lmp_Maximum_df, aes(x=temp,y=polyresiduals)) +
  geom_point() +
  stat_binmean(n = 10, color = "orange") +
  theme_bw() 

ggplot(lmp_Maximum_df, aes(x=temp,y=splineresiduals)) +
  geom_point() +
  stat_binmean(n = 11, color = "orange") +
  theme_bw() 

ggplot(lmp_Maximum_df, aes(x=temp,y=loessresiduals)) +
  geom_point() +
  stat_binmean(n = 11, color = "orange") +
  theme_bw() 


```

For the 7-th degree polynomial, it seems to perform better towards the ends of the distribution where we can see more clustering around zero. Towards the center of the distribution, we see a weaker performance.

For the cubic spline, we can see similar outcomes as the 7th degree polynomial, but the residuals that are positive are smaller in magnitude while the residuals that are negative are larger in magnitude.

The lowess distribution performs similarly as the previous two towards the high end of temperature, but towards the lower end of temperature and the center of the distribution, the residuals are marger in magnitude compared to the previous two models. However, in between these boundaries, the residuals seem to decrease in magnitude with this model compared to the others.































