---
title: "Problem Set 1"
author: "Sravan Ramawamy"
output:
  html_document: default
  pdf_document: default
---

https://github.com/SravanjR/bc-micro-methods/blob/main/psets/pset_auctions/Problem-Set-1.pdf

```{r setup, include=TRUE}
#Load Libraries 
library(MASS)
library(ggplot2)
library(evd)
```

```{r setupcont, include=TRUE}
#Load Bid Data
DataDir = paste(getwd(), "/bids1.csv",sep = "")
Data = read.csv(DataDir, header = FALSE)
bids = Data[,1]

```


## Question 1: Estimate Density of Bids

### Normal Distribution

```{r Q1, include=TRUE}

Normpdf = dnorm(bids, mean = mean(bids), sd = sd(bids))
Normpdf

```

### Gaussian Kernel

```{r gauss}

n = length(bids)
iq = IQR(bids)
iq
bw_plugin = .9*min(sd(bids),iq/1.34)*n^(-1/5)
bw_plugin
gk  <- density(bids,bw=bw_plugin, kernel = c("gaussian"))
gk
BAR1 <- with(gk, approxfun(x = x, y = y))
gDensity <- BAR1(bids)
gDensity

```

### Epanechikov Kernel

```{r epanechikov}

ek <- density(bids,bw=bw_plugin, kernel = c("epanechnikov"))
ek
BAR2 <- with(ek, approxfun(x = x, y = y))
eDensity1 <- BAR2(bids)
eDensity1

```

## Question 2: Least-Squares cross-validation for epanechikov kernel


```{r}
set.seed(1)
ek2 <- density(bids,bw="ucv", kernel = c("epanechnikov"))
ek2
BAR3 <- with(ek2, approxfun(x = x, y = y))
eDensity2 <- BAR3(bids)
eDensity2
```

## Question 3: Plot of Estimated Density Functions over Bid Histogram

```{r Q3}

data <- as.data.frame(bids)

ggplot(data,aes(bids))  + geom_histogram(aes(y = stat(density)), bins = 50) +
  geom_line(stat="density",bw=bw_plugin, col = 'red') + 
  geom_line(stat="density", bw=bw_plugin,lwd = 1, col = 'Orange', kernel = c("epanechnikov")) +
  geom_line(stat="density", bw="ucv",lwd = 1, col = 'blue', kernel = c("epanechnikov")) +
    stat_function(fun = dnorm, args = list(mean = mean(bids), sd = sd(bids)))

```

While the tails are somewhat similar, the the density function consisting of the epanechnikov and the plug-in estimate appear to fit the data best. This line is given by the orange curve above.

## Question 4: GPV and the cross-validated Epanechnikov kernel to recover the valuation implied for each bid

```{r}
BAR <- with(ek2, approxfun(x = x, y = y))
eDensity <- BAR(bids)

#Calculate CDF Values
f<- approxfun(ek2$x, ek2$y, yleft=0, yright=0)

cdf <- vector()
for(i in 1:length(bids)){
    temp <- integrate(f, lower = -Inf, bids[i])
    cdf[i] <- temp$value
}

value <- bids + cdf/(2*eDensity)
value
```

## Question 5: Estimate the Distribution of Value using another Epanechikov Kernel using the Plug-in Bandwidth

```{r}
n = length(value)
iq = IQR(value)
bw_plugin = .9*min(sd(value),iq/1.34)*n^(-1/5)
valDist  <- density(value,bw=bw_plugin, kernel = c("epanechnikov"))
valDist
```

## Question 6: Guess the Distribution that the Valuations were generated with

```{r}
m1 <- dgev(value, 2.25, 1.2, 0)

data2 <- as.data.frame(value)

ggplot(data,aes(value))  + geom_histogram(aes(y = stat(density)), bins = 50) +
geom_line(stat="density", bw=bw_plugin,lwd = 1, col = 'Orange', kernel = c("epanechnikov")) +
geom_line(aes(y = m1))
```

The Valuations seem to have been generated from a Generalized Extreme Value distribution. The Location parameter is around 2.25, the scale parameter is around 1.2 and the shape parameter is about 0 which suggests that the distribution of bidding valuations is specifically a Gumbel/Type 1 Extreme value Distribution.



